Impact of asymmetric tail-dependence in extreme climatic events on extinction risk in population ecology

### 2019-03-20 MainText.Rmd
add guidelines for intro and results

We explored the potential of asymmetric tail-dependence in copula 
for understanding the extinction risk of populations towards extreme climatic event in ecology
by addressing the following specific questions: (Q1) Does nature of tail dependence in vital rates
of a single-patch age-structured population model affect extinction risk of that population differently?
We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
(Q2) How does asymmetric tail-dependent noise across multiple patches influence extinction 
risk of a nonlinear metapopulation model having density dependence? We hypothesized the general finding 
from density-independent metapopulation model [BIVAN paper results of Lewinton-Cohen model]
(higher extinction risk if noise across patches were 
left-tail dependent than the one if noise came from right-tail dependent) holds true only in
the under-compensatory regime, but should be reverse if the model shows over-compensatory dynamics. 
(Q3) We further hypothesize that when there is no dispersal, the above-said switching should occur
near the critical parameter value where the deterministic model shows transition from 
under-compensatory to over-compensatory dynamics.

@@

@@

 <!--FIG : age structure model results : linear, nonlinear-->

\begin{figure*}[!h]
\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models, fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000. \label{fig_extrisk_agestr}}
\end{center}
\end{figure*}

@@

<!--Summary tables for 6 nonlinear models used-->
```{r read_tab_model_summary,echo=F, results='asis',message=F}
library(kableExtra)
library(dplyr)
dt<-readRDS("./Results/tab_model_summary.RDS")
knitr::kable(dt, escape=F,
             format="latex",digits=3, align="c",table.env='table*',
             caption = "Summary of six density dependence metapopulation models considered at each patch; last three columns indicated equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}^{i}$ : transition point from stable to unstable equilibrium of deterministic models), respectively. Superscript $i=1,2,...,n$ indicates states for $i^{th}$ patch.\\label{tab_model_summary}",
             booktabs=T, linesep = "\\addlinespace"
             )%>%column_spec(c(1,3,5,6),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
```

>>>>

<!--Background for extinction risk in ecological model-->

<!--How extinction risk is related to extreme events : past studies-->

<!--How copula could be relevent in estimating extreme event's influence on extinction risk : introduce Taildep. concept-->

<!--Tail dependence of copula and recent result from BIVAN showing LTdep with higher extrisk than UTdep, but 
not explored age structured or in metapopulation context with density dependence effect-->

<!--our objective ?-->
We explored the potential of asymmetric tail-dependence in copula 
for understanding the extinction risk of populations towards extreme climatic event in ecology
by addressing the following specific questions: (Q1) Does nature of tail dependence in vital rates
of a single-patch age-structured population model affect extinction risk of that population differently?
We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
(Q2) How does asymmetric tail-dependent noise across multiple patches influence extinction 
risk of a nonlinear metapopulation model having density dependence? We hypothesized the general finding 
from density-independent metapopulation model [BIVAN paper results of Lewinton-Cohen model]
(higher extinction risk if noise across patches were 
left-tail dependent than the one if noise came from right-tail dependent) holds true only in
the under-compensatory regime, but should be reverse if the model shows over-compensatory dynamics. 
(Q3) We further hypothesize that when there is no dispersal, the above-said switching should occur
near the critical parameter value where the deterministic model shows transition from 
under-compensatory to over-compensatory dynamics.

@@

<!--Q1 : age structured results : linear and nonlinear model with asymmetric copula-->

<!--Q1 : age structured results : linear and nonlinear model with extreme tail dep. copula-->

<!--Q2 : metapopulation results : density dependence with extreme tail dep. copula-->

<!--Q3 : metapopulation results : switching near rc for d=0-->

@@

 <!--FIG : age structure model results : linear, nonlinear-->
\begin{figure*}[!h]
\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models, fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000. \label{fig_extrisk_agestr}}
\end{center}
\end{figure*}

@@

<!--Summary tables for 6 nonlinear models used-->
```{r read_tab_model_summary,echo=F, results='asis',message=F}
library(kableExtra)
library(dplyr)
dt<-readRDS("./Results/tab_model_summary.RDS")
knitr::kable(dt, escape=F,
             format="latex",digits=3, align="c",table.env='table*',
             caption = "Summary of six density dependence metapopulation models considered for each patch; last three columns indicated equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}^{i}$ : transition point from stable to unstable equilibrium of deterministic models), respectively. Superscript $i=1,2,...,n$ indicates states for $i^{th}$ patch.\\label{tab_model_summary}",
             booktabs=T, linesep = "\\addlinespace"
             )%>%column_spec(c(1,3,5,6),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
```
